---
title: "PSQF 6250 - Assignment 1"
author: "Geraldo B. Padilla F."
date: "04-02-2022"
output: html_document
---

# Question 1

## **Research interests**:

- Multilevel modeling
- Computational analysis
- Quantitative methodologies
- Generalized linear models

### Hiperlinks to main R project website

Information about R project can be found here: <https://www.r-project.org/>. Or directly pressing [here](https://www.r-project.org/).

#### Question 4
```{r summ_iris}
summary(iris)
```

##### Question 5

| Descriptives | Sepal.Length | Sepal.Width | Petal.Length | Petal.Width|
| :---: | :---: | :---: | :---: | :---: | :---:|
| Min. | 4.300 | 2.000 | 1.000 | 0.100 |
| Median | 5.800 | 3.000 | 4.350 |  1.300|
| Max. | 7.900 | 4.400 | 6.900 | 2.500 |

###### Question 6

```{r hist_sepal, echo=FALSE}
hist(iris$Sepal.Length)
```

####### Questions 7

My favorites places to take vacations:

1. Bogota, Colombia
   + To visit the historical downtown
   + To drink coffee at the street
   + To climb up the Monserrate by cable car
2. Barcelona, Spain
   * To visit the Catalunya Square in front of the La Sagrada Familia
   * Walking through the Las Ramblas
   * To visit the Gaudí Park
3. Paris, France
   + To visit the Eiffel Tower
   + To go to the Louvre museum
   + Walking along the Senna river

######## Question 8

The next step in our analysis is exploring some correlations in the data set First, the relationship between the length and width of the sepals, which is moderately negative with a magnitude of `r round(cor(iris$Sepal.Length, iris$Sepal.Width),2)`. On the other hand, between the length and width of the petals there is a strong  positive correlation ( r = `r round(cor(iris$Petal.Length, iris$Petal.Width), 2)`).

######## Question 9

```{r plot_sepal, echo=FALSE}
plot(iris$Sepal.Length, iris$Sepal.Width)
```

