---
title: "Assignment 4"
subtitle: 'Inferential Statistics and Turning Tables into Graphics with R'
author: "Mubarak Mojoyinola, Geraldo Padilla"
date: "08-04-2022"
output:
  html_document:
    theme: simplex
    highlight: textmate
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(include = FALSE)
```

```{r packages, message=FALSE}
library(tidyverse)
library(stringr)
library(lubridate)
```

## Part 1: Inferential Statistics with R

```{r dataset}
d1 <- read.csv('https://raw.githubusercontent.com/lebebr01/statthink/main/data-raw/cal-fire-9-10-2020.csv')
```

1. Using text processing, create three new variables in the data that represents:

- the year the fire started
- the month the fire started
- Create a new variable that represents the length of time a fire burned. This can be created by using the difftime() function within R. Look at the help page to try to figure out how this function can compute differences in two dates.

```{r q1, include=TRUE}
d1 <- d1 %>% 
  mutate(date_incident =ymd_hms(incident_date_created),
         date_update = ymd_hms(incident_date_last_update),
         year = year(date_incident),
         month = month(date_incident),
         burn_days = as.numeric(difftime(date_update, date_incident, units = c("days"))))
```

2. Are there any data points that are extreme values or outliers that you feel should be removed from this analysis? Discuss briefly, why you feel these values may impact the analysis and should therefore be removed. Be as specific as you can why any values should or should not be removed. If you identify data that are suspect, remove them from further analysis (i.e. use filter to remove the values).

```{r}
summary(d1)

d1 <- d1 %>%filter(burn_days >= 0)
```

Almost all variables have values within the expected boundaries, excepting by burn_days, the variable we created to represents the length of time a fire burned. In this variable, some of the values were negative, so we decided to filter the dataset and allowing values at or above 0 (which is more meaningful that the previous version with values below 0).

3. Fit two competing models to attempt to answer the 1st and 2nd research questions. Summarize briefly the results from the models with particular attention to answering the research question so that non-statistics/data science individuals could use the answer for their planning or decision making process. Note, consider carefully the best approach on how to include month and year in your models (i.e., continuous vs factor type variables). 1 pt

```{r}
model1 <- lm(incident_acres_burned ~ I(year-min(year)), data = d1)
summary(model1)
```

Based on model1, there is no statistically significant evidence that the number of acres burned has increased for fires in more recent years. In other words, year is not a useful variable to make decision about how to predict or estimate the amount of acres burned in a fire incident. 

```{r}
model2 <- lm(incident_acres_burned ~ year*factor(month), data = d1)
summary(model2)
```

The addition of month as a factor variable to the model explains some variation in the number of acres burned but this is not really helpful, especially when we compare the R-square of both models: model1(0.0003079), model2(0.01739) and the adjusted R-squared of model2 (0.003827) is considered.

In summary, neither year nor month are relevant variables in order to estimate the number of acres burned in a fire incident. We should continue searching for other factors involved.

3.1 Which model fits best or do you feel is the best model?

The two models do not have significant estimates, therefore, we will prefer the simpler model (model1).

4. Check assumptions for the models from #3. Does there appear to be problems with meeting statistical assumptions? Provide rationale for why or why not.

```{r, eval=FALSE}
plot(model1)
```

Exploring model1 assumptions, none of these are met. The model residuals are not normally distributed, the variance looks like heterogeneous, and it has some leverage points that could be affecting the model performance.

The situation is not different for model2. It also has problems with the assumptions, especially with homoscedasticity. In this plot, there is a trend in the residuals as the values of the explanatory variables increase.

```{r, include=TRUE, echo=FALSE, warning=FALSE, eval=FALSE}
plot(model2, 3)
```

5. Fit another model to attempt to answer the third research question. Summarize briefly the results from the model with particular attention to answering the research question so that non-statistics/data science individuals could use the answer for their planning or decision making process.

```{r}
model3 <- lm(incident_acres_burned ~ year+burn_days+factor(month), data = d1)
summary(model3)
```
Once burn_days has been added to the model, the results do not change significantly. Based on the R-squared and Adjusted R-squared, model3 performs even worst than the previous versions. In other words, analyzing the year, month and days when a fired incident has occurred, we have not found nothing statistically significant to estimate the amount of acres burned. Thus, the challenge is to founding additional information which helps to answer our questions. 

6. Finally, create a graphic that summarizes the results from your final model that you feel fits the data the best (i.e. this could be the model from #4 or #7). Discuss why you picked this model and describe why this figure does a good job of showing the takeaway message.

```{r, eval=FALSE}
ggplot(d1, aes(as.factor(month), incident_acres_burned))+
  geom_point(color = 'red')+
  facet_wrap(~year)+
  labs(x = 'Month', y = 'Acres burned')+
  theme_bw()
```

We have chosen the second model because it is simple and clear, and also because the third models is more complex and it does not add nothing meaningful to our analyses. We have also divided the graphs based on the different years when the fires occurred to show the relation between the amount of acres burned and months. In the different facets we can see that there is neither a clear nor a slight pattern in the data.
It appears that the acres burned and the months and years are not related.

## Part 2: Turning Tables into Graphics with R

1. Find a published table showing descriptive or inferential statistic results. Provide a copy of the table (a screen shot is fine for this purpose).

```{r}
image <- "https://github.com/Bladimir-PF/PSQF-6250-Computer-Packages-/blob/GLM/Vegetables%20and%20Fruits.PNG"
download.file(image, 'image.png')
file.show(image)
```





2. Turn the table from #1 above into a publication quality graphic using R. Ensure that this graphic conveys the same purpose as the original table. 1 pt

3. Create a different from the figure created in #3 above, and make an interactive graphic that attempts to convey the same message as the original table. Note, this figure should be an entirely different figure from that in #3. For example, if you created a bar chart in #3, create something other than a bar chart for this question. 1 pt

4. Briefly discuss whether you feel the graphs convey the message better or worse than the original table. Use specific examples from the table/graph and recommendations from the article in your discussion. Which figure do you feel does the best job in sharing the original purpose of the table? Be specific in your discussion. 1 pt

5. Take the data from the original table, to do this you may need to create an Excel file or use the function data.frame() to import the data from the original table. Use the kable() function and kableExtra package to create a reproducible table that looks as close as possible to the original table. 1 pt
